---
title: "Model Fitting"
author: "Alex Gichuki"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(tidymodels)
```


```{r data, include = FALSE}
hfi <- read_csv('https://www.openintro.org/data/csv/hfi.csv')

dim(hfi)
```

The hfi data set has 1458 rows and 123 variables. Each row represents a countries data for a specific year.

### Exploring Data for 2016

```{r year 2016, include=FALSE}
hfi_2016 <- filter(hfi, year == 2016)
```

Creating a scatter plot to evaluate the relationship between pf_score and pf_expression_control.
I plot Expression Control on the x axis since it is the predictor variable.
```{r scatter plot, include=TRUE}
ggplot(data = hfi_2016, aes(pf_expression_control, pf_score)) +
  geom_point(alpha = 0.5, size= 2) +
  labs(x = "Expression Control", y = "Score",
       title = "Scatter Plot for Expression Control and Score") 
  
```


It looks like there could be a linear relationship in the scatter plot and we could use a linear model to predict personal freedom score.

## Sum of Squared Residuals

```{r Sum of Squared Residuals}
# Using trial and Error

statsr::plot_ss(x = pf_expression_control, y = pf_score, data = hfi_2016, showSquares = TRUE)
```

It is not easy to get the correct least squares line using this method.
The least sum of squares I got was 278.852.

## The Linear Model

```{r Sum of Squared Residuals 2, echo=TRUE}
# Using lm function from base R to plot a regression line
m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)

tidy(m1)

```

From the code chunk above, the least squares regression line for the model is pf_score = 4.28 + 0.542(pf_expression_control)

For countries with a pf_expression_control of 0 (those with the largest amount of political pressure on media content), we expect their mean personal freedom score to be 4.28.

For every 1 unit increase in pf_expression_control, we expect a country’s mean personal freedom score to increase 0.542 units.

## Overall Model Fit
```{r Correlation Coefficient}
# Obtaining the correlation coefficient between expression control and score.

cor(hfi_2016$pf_expression_control, hfi_2016$pf_score)
```

Both variables have a correlation coefficient of 0.8450646 which indicates a strong positive correlation. An increase in expression control translates to an increase of 0.8450646 on the pf_score.


Now using R Squared to assess the model:
```{r R squared}
glance(m1)
```

R Squared for this model is 0.714. This indicates that 71.4% of the variation in the score variable (dependent variable) can be explained by expression control (independent variable).


Now creating a model for using pf_expression_control to predict hf_score:
```{r model 2}
m2 <- lm(hf_score ~ pf_expression_control, data = hfi_2016)

tidy(m2)
```

From the code chunk above, the least squares regression line equation for the model is hf_score = 5.05 + 0.368(pf_expression_control)

For countries with a pf_expression_control of 0 (those with the largest amount of political pressure on media content), we expect their mean human freedom score to be 5.05.

For every 1 unit increase in pf_expression_control, we expect a country’s mean human freedom score to increase 0.368 units.


## Prediction and Prediction Errors
Creating a scatter plot with a smooth line.

```{r Scatter plot with smooth line}
ggplot(data = hfi_2016, aes(pf_expression_control, pf_score)) +
  geom_point(alpha = 0.5, size= 2) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Expression Control", y = "Score",
       title = "Scatter Plot for Expression Control and Score")
```

## Model Diagnostics
Check linearity, nearly normal residuals, and constant variability to verify if the linear model is reliable:

```{r}
m1_aug <- augment(m1)
m1_aug
```

Plotting the residuals vs fitted (predicted) values:
Overlaying a red horizontal dashed line at 
 (to help us check whether the residuals are distributed around 0)
```{r}
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")

```

This plot does not show an apparent pattern which indicate there is a linear relationship between the two variables.

Checking for nearly normal residuals:

```{r}
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

